What's in a Domain? Anaylsis of URL Features

### 2022-09-29 paper/paper.tex
Paper development

\begin{abstract}
Many data science problems require processing log data derived from web pages, apis or other
internet traffice sources. URLs are one of the few universal pieces of information that describe
internet activity, hence URLs require effective processing for a wide variety of machine 
learning applications. While URLs are structurally rich, they are not subject
to universal principles of construction making fetaure engineering for internet data an ongoing challenge.

In this research we outline the key categories of information structure in domains and URLs.
We share an open source implementation of these ideas and demonstrate their utility across a range of 
internet traffic machine learning problems.
\end{abstract}

@@

Multiple data science and machine learning applications on internet data 
Write a literature review \cite{Vazhayil2018}
\cite{Canali2011}
\cite{Ayes2019}
\cite{Mamun2016}
\cite{Li2020}
\cite{Basnet2012}
\cite{Basnet2014}
\cite{Tupsamudre2019}
URLnet learns word and charcater embedings \cite{Le2018}
\cite{Baykan2009}
\cite{Kan2005}
\cite{Hernandez2012}
\cite{Chung2010}
\cite{Qiu2020}
\cite{Vaishnavi2021}
\cite{Xu2021}
\cite{Rhea2022}
\cite{Ma2009}
\cite{Verma2017}
\cite{Sountharrajan2020}
Classify Web pages based on URLs from the 4 Universities WebKB dataset. Comparing features derived from segmentation
of the URL to features derived from either link anchor text or the web page document itself \cite{Kan2004}.

@@

@@

\subsection{Contextual Categorisation}

Verity had returned the
highest score and converted that category into a set of labels for
classification across the four Tiers of iAB V2.0. In instances where the
classified category is from a higher position in the hierarchy, then the
lower level tiers are given the deepest category. For example, if we have
a URL that is classified as iABv2 category "Business and Finance"
(a Tier 1 category) then Tiers 2, 3 and 4 will also contain the same value.

Alternatively, if we have another URL that is classified as "Marketing and Advertising"
(a Tier3 category below "Business and Finance"), then the hierarchical labels
would be assigned as follows:

\begin{itemize}
    \item {\texttt{Tier 1}}: ["Business and Finance"]
    \item {\texttt{Tier 2}}: ["Business"]
    \item {\texttt{Tier 3}}: ["Marketing and Advertising"]
    \item {\texttt{Tier 4}}: ["Marketing and Advertising"]
\end{itemize}

The idea of this scheme is that every page will be assigned a categorisation across
all four levels in the hierarchy, but sometimes the category at a given Tier
will not be any more specific than the parent category. This approach ensures
that comparisons can always be made, and when we look at a deeper level in the hierarchy
we will have at least as many distinct categories as the parent.

\subsection{Metrics}

The focus of our analysis is to determine whether the performance of each creative,
in a set of contexts, remains consistent across the two time periods.
Within each time period we aggregate
the measured attention time over the impression level data to obtain multiple metrics
for the reliability of attention time within a given context.

The attention metrics are calculated for a set of impressions $I_p$ in a time period $p$.
We first define a baseline for a creative $c$ in period $p$ as:
${}^c\mathcal{A}_p$. This is defined in equation \ref{campaign_baseline}.

@@

Where the function $A(i)$ returns the attention measured on impression $i$ and
$\mathbf{1}^c(i)$ is an indicator function that
tests whether impression $i$ belongs to creative $c$, as shown in equation \ref{indicator}.

\begin{equation}
\label{indicator}
  \mathbf{1}^{c}(i):=
  \begin{cases}1~&{\text{ if }}~i\in c~,\\0~&{\text{ if }}~i\notin c~.\end{cases}
\end{equation}

Similarly, we denote the baseline attention achieved by ads on pages classified
as belonging to category $t$ in period $p$ as: ${}_t\mathcal{A}_p$. Defined in
equation \ref{topic_baseline}. Note, we use the notation $t$ for categories as they
are equivalent to topics, and this distinguishes them from the creatives.

\begin{equation}
\label{topic_baseline}
{}_t\mathcal{A}_p =  \frac{ \sum_{i \in I_p} \mathbf{1}_t(i) A(i) }{ \sum_{i \in I_p} \mathbf{1}_t(i) }
\end{equation}

Where $\mathbf{1}_t(i)$ is an indicator function that
tests whether impression $i$ was shown on a page classified as category $t$,
as show in in equation \ref{indicator2}.

\begin{equation}
\label{indicator2}
  \mathbf{1}_{t}(i):=
  \begin{cases}1~&{ i\text{ categorised as }}~t~,\\0~&{\text{ otherwise }}.\end{cases}
\end{equation}

We denote the mean attention performance for a creative $c$, on media of category $t$
within period $p$ as ${}^c_t\mathcal{A}_p$.
This is defined as shown in Equation \ref{mean_attention}.

\begin{equation}
\label{mean_attention}
{}^c_t\mathcal{A}_p =  \frac{ \sum_{i \in I_p} \mathbf{1}^c_t(i) A(i) }{ \sum_{i \in I_p} \mathbf{1}^c_t(i) }
\end{equation}

Where we simply apply a indicator function $\mathbf{1}^c_t(i)$ that is a composite
of Equations \ref{indicator} and \ref{indicator2} to filter for inventory that
belongs both to creative $c$ and category $t$. These equations should illustrate that we
make our evaluations by looking at the expected attention time across difference slices
of the inventory.

\subsection{Analysis}

Our analysis proceeds by ranking all topics that each creative $c$ was exposed
to according to the value of ${}^c_t\mathcal{A}_1$.
We then look at the same creatives in period 2 and
evaluate ${}^c_t\mathcal{A}_2$ for the top three ranking categories from the first period,
as well as the bottom ranking category.
This approach necessitates that each creative be exposed to a minimum
of these same 4 categories in both periods.
In the results that will follow we report the number
of records in each trial, reflecting the number of aggregated records
that fulfills these requirements.

Once we have the datasets in each period that meet these requirements we use multiple metrics
to evaluate the impact of the contextual categories. Our primary metric is the percentage
of instances in which the category specific performance ${}^c_t\mathcal{A}_2$ is above the
creative's baseline expected performance ${}^c\mathcal{A}_1$.
This is shown as \textbf{\% Over} (the baseline) in the tables of results.

In addition we look at the mean difference between ${}^c_t\mathcal{A}_2$ and ${}^c\mathcal{A}_1$
over all creatives in the experiment. This is denoted \boldmath$\mu\delta$ (or mean delta) in
the tables and defined in Equation \ref{delta}

\begin{equation}
\label{delta}
\mu\delta = \frac{ \sum_{c \in \mathcal{C}} {}^c_t\mathcal{A}_2 - {}^c\mathcal{A}_1 }{ \mathcal{C} }
\end{equation}

The mean delta illustrates an expectation of the difference between attention time performance
and the creative baseline, using a specific contextual targeting strategy. In addition, we show
that difference expressed as a mean percentage (against the baseline) in the column
$\mathbf{\mu\%\delta}$. This can be interpreted as the expected percentage increase/decrease
relative to the creative's baseline attention, when using that contextual targeting strategy.

@@

We conduct two variations of the experiment, each one looking at different
criteria for evaluating performance on new inventory.
These are referred to as the
\emph{No URL Repeats} and the \emph{No Domain Repeats} experiments.

\subsection{No URL Repeats}

In Table \ref{tab:nourlrepeats} we see the results of looking at performance on novel
URLs for a creative, where the iAB V2.0 contextual categories of those URLs were present
in either the top three or last ranking position in the initial period.

\begin{table}
\caption{Attention Performance by Category Rank on Novel URLs}
\label{tab:nourlrepeats}
\begin{tabular}{|l|r|l|r|r|r|}
\toprule
Level   &Count    &Rank &\% Over  &$\mu\delta$  &$\mu\%\delta$        \\
\midrule
Tier1   &206      &1st  &100.0\%    &4.5         &93.2\%     \\
        &         &2nd  &96.6\%     &2.6         &49.8\%     \\
        &         &3rd  &80.6\%     &1.4         &28.7\%     \\
        &         &Last &0.0\%      &-3.7        &-53.3\%     \\
Tier2   &210      &1st  &99.5\%     &6.3         &132.0\%     \\
        &         &2nd  &95.7\%     &3.7         &68.6\%     \\
        &         &3rd  &80.0\%     &2.4         &41.2\%     \\
        &         &Last &0.0\%      &-4.1        &-59.2\%     \\
Tier3   &210      &1st  &99.5\%     &6.4         &135.7\%     \\
        &         &2nd  &95.7\%     &3.8         &69.5\%     \\
        &         &3rd  &79.5\%     &2.4         &41.8\%     \\
        &         &Last &0.0\%      &-4.1        &-59.2\%     \\
Tier4   &210      &1st  &99.5\%     &6.4         &135.7\%     \\
        &         &2nd  &95.7\%     &3.8         &69.5\%     \\
        &         &3rd  &79.5\%     &2.4         &41.8\%     \\
        &         &Last &0.0\%      &-4.1        &-59.2\%     \\
\bottomrule
\end{tabular}
\end{table}

We see that, in general, targeting pages using categories deeper into the hierarchy
improves the expected increase in attention through contextual targeting.
However, the effect appears to plateau at Tier3, likely due to the small proportion of inventory
that is uniquely classified at a Tier 4 level.

\subsection{No Domain Repeats}
We repeat the analysis restricting ourselves to impressions in the second period
that involve URLS for which the domain was not present in the URLs for the first period.
The results are shown in Table \ref{tab:nodomainrepeats}.
We note that the number of records has dropped by
more than half in this experiment, due to the difficulty of finding domains that were
not present in the first period.

\begin{table}
\caption{Attention Performance by Category Rank on Novel Domains}
\label{tab:nodomainrepeats}
\begin{tabular}{|l|r|l|r|r|r|}
\toprule
Level   &Count  &Rank &\% Over  &$\mu\delta$  &$\mu\%\delta$        \\
\midrule
Tier1   &86       &1st  &98.8\%     &5.2         &144.3\%     \\
        &         &2nd  &83.7\%     &2.3         &58.8\%     \\
        &         &3rd  &60.5\%     &0.4         &26.9\%     \\
        &         &Last &0.0\%      &-4.1        &-64.5\%     \\
Tier2   &83       &1st  &96.4\%     &6.4         &159.1\%     \\
        &         &2nd  &84.3\%     &2.6         &71.0\%     \\
        &         &3rd  &59.0\%     &0.7         &35.8\%     \\
        &         &Last &0.0\%      &-4.5        &-70.0\%     \\
Tier3   &82       &1st  &96.3\%     &6.5         &161.3\%     \\
        &         &2nd  &84.2\%     &2.7         &71.2\%     \\
        &         &3rd  &56.1\%     &0.6         &34.7\%     \\
        &         &Last &0.0\%      &-4.6        &-69.9\%     \\
Tier4   &82       &1st  &96.3\%     &6.5         &161.3\%     \\
        &         &2nd  &84.2\%     &2.7         &71.6\%     \\
        &         &3rd  &56.1\%     &0.6         &34.7\%     \\
        &         &Last &0.0\%      &-4.6        &-69.9\%     \\
\bottomrule
\end{tabular}
\end{table}

The restriction to unique domains in the second period has lifted the expected improvement
of targeting the top performing category (Tier 3) from 135% to 161%. However, this improvement
only occurs with the category in ranking position one. For the other ranking positions
the results are mixed, even though the percentage amount is sometimes higher, the
absolute amount can be lower, due to the fact that the baseline is calculated on
the impressions that qualify for inclusion.

These result seems counterintuitive at first glance. Our current hypothesis about why
this is the case is that it relates to the specificity of the content on a given domain.
The \emph{no domain repeats} study is biased towards data on less common domains.
These may have a tendency to be more subject specific.
These ideas need to be tested in follow up research.

Additionally, we note that the percentage of instances where the category targeting
delivers an improvement over the baseline drops when we look at unique domains.
This drop is approximately 3% for rank 1 categories, but by the time we arrive at
the 3rd ranked category it has dropped from 80% to 56% at Tier 4, meaning that
targeting new domains with the 3rd ranked category offers a marginally better than
random chance of delivering improved attention.

>>>>

\begin{abstract}
Many data science problems require processing log data derived from web pages, apis or other
internet traffic sources. URLs are one of the few ubiquitous data fields that describe
internet activity, hence they require effective processing for a wide variety of machine 
learning applications. While URLs are structurally rich, they are not subject
to universal principles of construction making fetaure engineering for internet data an ongoing challenge.

In this research we outline the key categories of information structure in domains and URLs.
We share an open source implementation of these ideas and demonstrate their utility across a range of 
internet traffic machine learning problems. Python package available at https://pypi.org/project/url2features
\end{abstract}

@@

Multiple data science and machine learning applications on internet data 
Write a literature review \cite{Vazhayil2018}
\cite{Canali2011}
\cite{Ayes2019}
\cite{Mamun2016}
\cite{Li2020}
\cite{Basnet2012}
\cite{Basnet2014}
\cite{Tupsamudre2019}
URLnet learns word and charcater embedings \cite{Le2018}
\cite{Baykan2009}
\cite{Kan2005}
\cite{Hernandez2012}
\cite{Chung2010}
\cite{Qiu2020}
\cite{Vaishnavi2021}
\cite{Xu2021}
\cite{Rhea2022}
\cite{Ma2009}
\cite{Verma2017}
\cite{Sountharrajan2020}

\cite{Murthy2022}
Classify Web pages based on URLs from the 4 Universities WebKB dataset. Comparing features derived from segmentation
of the URL to features derived from either link anchor text or the web page document itself \cite{Kan2004}.

@@

\begin{table*}
\caption{URL Features}
\label{tab:features}
\begin{tabular}{|l|l|l|l|}
\toprule
Group         &Feature              &Type        &Definition  \\
\midrule
Protocol      &protocol\_name        &Category    &The exact protocol name    \\
              &protocol\_type        &Category    &Protocol categorised by purpose    \\
              &protocol\_exists      &Boolean     &Flag indicating presence of protocol   \\
\midrule
Host          &host\_is\_ip           &Boolean     &Flag indicating if the host is an IP address    \\
              &ext\_type             &Category    &TLD extension categorised by purpose    \\
              &ext\_freq             &Boolean     &TLD extension frequency in internet traffic   \\
\midrule
Path          &path\_depth           &Numeric     &The number of forward slashes    \\
              &path\_words           &Numeric     &Count of distinct words    \\
\midrule
File          &file\_ext             &Category    &The exact file extension e.g. "exe" or "php"    \\
              &file\_type            &Category    &File extension categorised by function \& purpose    \\
              &file\_ext\_exists      &Boolean     &Flag indicating presence of protocol   \\
\midrule
Params        &name                 &Category    &The exact protocol name    \\
              &type                 &Category    &Protocol categorised by purpose    \\
              &exists               &Boolean     &Flag indicating presence of protocol   \\
\bottomrule
\end{tabular}
\end{table*}

These URL feature functions are all available in the open source python package \emph{url2features}
available at https://pypi.org/project/url2features with links to usage documentation and source
code.

@@

\subsection{Feature Importance}

@@

@@
